The Implications of Local Fluctuations in the Galactic Midplane for Dynamical Analysis in the \textit{Gaia} Era

### 2019-07-26 paper/ms.tex
Clean up comments throughout

\textbf{The true Galactic potential is not strictly axisymmetric, beyond even
the typically-quoted gas clouds, spiral arms, and bar(s). The presence of disk
oscillations has been known
%mm for some time, observed in 1957 for
% H~% \textsc{i} in the outer disk \citep{1957AJ.....62...93K}.
    since its observation in H~{sc i} by \citet{1957AJ.....62...93K}.
The presence of a North-South
asymmetry in stellar number density and velocity in the solar neighborhood has
been found in
% mm [should these multiple acronyms be expanded?]
SDSS data \citep{2012ApJ...750L..41W}, and in velocity in the
LAMOST survey \citep{2013ApJ...777L...5C, 2015ApJ...801..105X}, and RAVE data
\citep{2013MNRAS.436..101W}. Beyond the solar neighborhood, low-latitude
overdensities have been observed, e.g., the Monoceros Ring
\citep{2002ApJ...569..245N, 2003MNRAS.340L..21I, 2003ApJ...594L.119C,
2014ApJ...791....9S, 2016ApJ...825..140M}, an overdensity in the direction of
the Triangulum and Andromeda galaxies \citep{2007ApJ...668L.123M,
2014ApJ...793...62S, 2015ApJ...801..105X, 2015MNRAS.452..676P} and A13
\citep{2010ApJ...722..750S, 2017ApJ...844...74L}, and in the direction of the
Galactic Center \citep{2014Natur.509..342F}. There is evidence that these
stellar populations originated from the disk \citep{2015MNRAS.452..676P, 2018ApJ...854...47S,
2018Natur.555..334B}. See, e.g., \citet{2018MNRAS.481..286L} for a discussion
of these observations.}

@@

% mm Small-scale
     Such
variations in the density of gas and stars cause the local
midplane position to vary as a function of radius and azimuth. Stars far from
the Sun have a local midplane that differs from our local midplane
extrapolated onto their position. Converting positions and velocities of more
distant stars from a heliocentric to a Galactocentric coordinate system thus
introduces a systematic bias in the $z$ coordinate. We show that this bias
induces non-Gaussian errors in the actions computed for these stars. The
further from the solar neighborhood the target star is, the more likely the
mismatch will result in large systematic uncertainty, especially in the
vertical action $J_z$. A similar argument applies to any remaining uncertainty
in measurements of the Galactocentric radius of the Sun, and to variations in
the LSR.

@@

%mm [moved para up as it directly follows previous para]
This example uses an offset in $z$, but analagous effects occur from offsets
in other coordinates, such as the distance to the Galactic center or the LSR.
A similar effect in which actions gain time-dependence due to a time-varying
potential was pointed out by \citet{2015A&A...584A.120B}.

\textbf{We would like to emphasize that in reality one only ever measures the
position and velocity of a star at a single instance in time, and not over any
significant portion of its orbit. However, if more than one star on the same
orbit
%mm can be identified, then it might be possible 
     could be identified, then it would be possible
to constrain the disk potential in analogy to how stellar streams are
used to constrain the Galactic  potential
\citep[e.g.][]{1999ApJ...512L.109J, 2001ApJ...551..294I, 
2004ApJ...610L..97H, 2015ApJ...801...98S, 2015ApJ...803...80K,
2019MNRAS.483.1427L}. However, identifying the remnants of a disrupted open
cluster (the inner disk equivalent of a halo stellar stream) is a difficult
challenge. One
%mm would want to use the same constants of motion that we discuss are
%suscept to significant errors. On the other hand, one could use
    possibility would be to use the actions, but as we discuss in this
    work they are subject to substantial errors.  Another possibility would
    be to use
chemical tagging to identify a disrupted open cluster \citep{2002ARA&A..40..487F}, but
%mm it
    this
may be difficult due to the large number of stars that are chemically
indistinguishable
%mm but are
    though
presumably born in different clusters
\citep{2018ApJ...853..198N}. An alternative technique that would deliver some
of the benefits of finding a disk stream may lie in measuring the
accelerations of stars \citep{2018arXiv181207581S}.}

%mm This example uses an offset in $z$, but analagous effects occur from offsets
%in other coordinates, such as the distance to the Galactic center or the LSR.
%A similar effect in which actions gain time-dependence due to a time-varying
%potential was pointed out by \citet{2015A&A...584A.120B}.

@@

We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
but for four succcessive timesteps of the live N-body simulation of a
Sagittarius encounter \citep{2018MNRAS.481..286L}. Again we have subtracted a
best fit curve of the form given in Equation~(\ref{eq:fit-curve}), with the
values at times
% mm  [line breaks are needed]
%$(2.0, 4.0, 6.0, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -210,
%-390)\,\pc$, $B = (0.074, 0.039, -36, -57)\,\text{deg}$, and $C = (2.6,-7.8,
%-65, -53)\,\pc$. The middle $90\%$ of midplane values across the solar circle
%spans mm $(49, 62, 140, 120)\,\pc$.
    (2.0, 4.0, 6.0, 6.9)~Gyr being $A$ = (9.5, 2.5, -210, -390)~pc,
    $B$ = (0.074, 0.039, -36, -57)~deg, and $C$ = (2.6,-7.8, -65,
    -53)~pc. The middle $90\%$ of midplane values across the solar
    circle spans (49, 62, 140, 120)~pc.

>>>>

\textbf{The true Galactic potential is not strictly axisymmetric, beyond even
the typically-quoted gas clouds, spiral arms, and bar(s). The presence of disk
oscillations has been known
    since its observation in H~{\sc i} by \citet{1957AJ.....62...93K}.
The presence of a North-South
asymmetry in stellar number density and velocity in the solar neighborhood has
been found in
SDSS data \citep{2012ApJ...750L..41W}, and in velocity in the
LAMOST survey \citep{2013ApJ...777L...5C, 2015ApJ...801..105X}, and RAVE data
\citep{2013MNRAS.436..101W}. Beyond the solar neighborhood, low-latitude
overdensities have been observed, e.g., the Monoceros Ring
\citep{2002ApJ...569..245N, 2003MNRAS.340L..21I, 2003ApJ...594L.119C,
2014ApJ...791....9S, 2016ApJ...825..140M}, an overdensity in the direction of
the Triangulum and Andromeda galaxies \citep{2007ApJ...668L.123M,
2014ApJ...793...62S, 2015ApJ...801..105X, 2015MNRAS.452..676P} and A13
\citep{2010ApJ...722..750S, 2017ApJ...844...74L}, and in the direction of the
Galactic Center \citep{2014Natur.509..342F}. There is evidence that these
stellar populations originated from the disk \citep{2015MNRAS.452..676P, 2018ApJ...854...47S,
2018Natur.555..334B}. See, e.g., \citet{2018MNRAS.481..286L} for a discussion
of these observations.}

@@

Such
variations in the density of gas and stars cause the local
midplane position to vary as a function of radius and azimuth. Stars far from
the Sun have a local midplane that differs from our local midplane
extrapolated onto their position. Converting positions and velocities of more
distant stars from a heliocentric to a Galactocentric coordinate system thus
introduces a systematic bias in the $z$ coordinate. We show that this bias
induces non-Gaussian errors in the actions computed for these stars. The
further from the solar neighborhood the target star is, the more likely the
mismatch will result in large systematic uncertainty, especially in the
vertical action $J_z$. A similar argument applies to any remaining uncertainty
in measurements of the Galactocentric radius of the Sun, and to variations in
the LSR.

@@

This example uses an offset in $z$, but analagous effects occur from offsets
in other coordinates, such as the distance to the Galactic center or the LSR.
A similar effect in which actions gain time-dependence due to a time-varying
potential was pointed out by \citet{2015A&A...584A.120B}.

\textbf{We would like to emphasize that in reality one only ever measures the
position and velocity of a star at a single instance in time, and not over any
significant portion of its orbit. However, if more than one star on the same
orbit
     could be identified, then it would be possible
to constrain the disk potential in analogy to how stellar streams are
used to constrain the Galactic  potential
\citep[e.g.][]{1999ApJ...512L.109J, 2001ApJ...551..294I, 
2004ApJ...610L..97H, 2015ApJ...801...98S, 2015ApJ...803...80K,
2019MNRAS.483.1427L}. However, identifying the remnants of a disrupted open
cluster (the inner disk equivalent of a halo stellar stream) is a difficult
challenge. One
    possibility would be to use the stellar actions or other conserved quantities, but as we discuss in this
    work they are subject to substantial errors.  Another possibility would
    be to use
chemical tagging to identify a disrupted open cluster \citep{2002ARA&A..40..487F}, but
    this
may be difficult due to the large number of stars that are chemically
indistinguishable
    though
presumably born in different clusters
\citep{2018ApJ...853..198N}. An alternative technique that would deliver some
of the benefits of finding a disk stream may lie in measuring the
accelerations of stars \citep{2018arXiv181207581S}.}

@@

We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
but for four succcessive timesteps of the live N-body simulation of a
Sagittarius encounter \citep{2018MNRAS.481..286L}. Again we have subtracted a
best fit curve of the form given in Equation~(\ref{eq:fit-curve}), with the
values at times
    (2.0, 4.0, 6.0, 6.9)~Gyr being $A$ = (9.5, 2.5, -210, -390)~pc,
    $B$ = (0.074, 0.039, -36, -57)~deg, and $C$ = (2.6,-7.8, -65,
    -53)~pc. The middle $90\%$ of midplane values across the solar
    circle spans (49, 62, 140, 120)~pc.
